The Galactic Chemical Evolution of Carbon: Implications for Stellar Nucleosynthesis

### 2023-07-11 carbon_paper/tex/main.tex
new draft

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

% Acronyms
\newcommand{\agb}{AGB}
\newcommand{\apogee}{APOGEE}
\newcommand{\cc}{CCSNe}
\newcommand{\gce}{GCE}
\newcommand{\ia}{SNe Ia}
\newcommand{\imf}{IMF}
\newcommand{\sfh}{SFH}
\newcommand{\ssp}{SSP}

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. A. Boyea et. al.]{
Daniel A. Boyea,$^{1}$\thanks{E-mail: boyea.2@osu.edu}
James W. Johnson,$^{1}$
Third Author$^{2,3}$
and Fourth Author$^{3}$
\\
% List of institutions
$^{1}$Department of Astronomy, the Ohio State University, 191 W. Woodruff, Columbus, OH 43210, USA
}

@@

% Abstract of the paper
\begin{abstract}
% context
C is an important element across astronomy. However, its origin remains poorly understood. 
% 
As stellar yields are highly uncertain, we aim to constrain the stellar yields of C through multi-zone Galactic chemical evolution models. 
% results
We use \apogee\ subgiants to constrain the model. 
% 
We find that \caafe\ is an emirical estimate of the delayed C sources, enabling us to estimate that \agb\ stars and \cc\ produce about 20\% and 80\% of C, respectively.  
The \caah\ trend instead represents the equilibrium abundances of C and Mg. 
We use the \caah\ trend to estimate the CCSNe C/Mg yield, determining that  $\Ycc/\Yoc = EQUATION$ when including \agb\ C. 
% misc
Our models are relatively independent of uniform scaling of yields and outflows, alternate \sfh{}s. 
However, the stars which contribute to \agb\ C production and the \ia\ delay time distribution of Fe contribute uncertainties to our conclusions. 
% gas phase
While reliable gas-phase and low-metallicity measurments of C are challenging, we find that our model and a singlezone model with our recommended yields replicate the broad trends of \caah across different environments and metallicities.

@@

One of the primary uncertainties of GCE models are nucleosynthetic yields. Yield predictions -- the amounts of each chemical element stars produce --
are shaped by poorly understood processes, including mass loss, nuclear reaction rates, rotational mixing, convection, and explodability \citep{romano+10,KL14,ventura+13, LC18, emily+21}.
To better understand where C comes from and how it evolves, our aim is to use a combination of \apogee\ observations and multi-zone models to develop estimates of the required C yields to reproduce galactic trends.
\cite{james+23} examined similar \gce{} models of N (which is closely related to C), finding that trends in N and O\footnotemark{} are explained by the metallicity dependence of N/O yields. \citet{james+23} determine that \agb\ N abundances roughly dependence on metallicity (i.e. $y_{\rm N}/y_{\rm O} \propto Z$). 
Here, we extend their models to C, deriving similar constraints on C/Mg yields. We assess which yield prescriptions reproduce Galactic abundance trends while investigating the impact of \gce{} model assumptions, such as \sfh{} and outflow mass loading.

@@

\begin{figure*}
    \centering
    \includegraphics{subgiants.pdf}
    \caption[Subgiant Abundances]{The [C/Mg] ratio against [Mg/H] (top) and [Mg/Fe] (bottom) for the \citetjack~sample of \apogee{} subgiants. On the top, we plot high and low-$\alpha$ stars in blue and orange, using the separation defined in Equation \ref{eq:high_alpha} (the high and low-$\alpha$ stars are named for their high or low $\alpha$-element to Fe ratios, or in this case, Mg/Fe). On the bottom, we colour-code stars according to their [Mg/H] abundance.} \label{fig:subgiants}
\end{figure*}

@@

\begin{table}
	\centering
    \caption[Fiducial Model Yields]{Yields for the fiducial model (in units of \ssp~birth mass). See Section \ref{sec:agb} for the definition of \cxi.}
	\label{tab:fiducial_mod}

@@

Here, we assume a simple \textit{one-zone} chemical evolution model \cite[e.g.][]{tinsley80, pagel09, matteucci21}.  Newly produced metals are homogeneously and instantaneously mixed, so this model evolves as a single, spatially-uniform chemical zone.
We define $M_{X}$ to be the mass of element $X$ in the gas-phase, $\dot{M}_\star$ to be the star formation rate (in $\Mo\,{\rm yr}^{-1}$), and $\eta$ to be the mass loading factor $\eta\equiv\dot{M}_{\rm outflow}/\dot{M}_\star$ (representing the strength of outflows). A \ssp\ returns a fraction $r$ of their birth mass to the interstellar medium, due to ejected stellar envelopes.%
\footnote{$r\approx0.4$ for a \citealt{kroupa01} \imf.}
Here, we will use $\alpha$ as a representative $\alpha$-element, such as O or Mg. 
Given the \imf-averaged yield $y_\alpha$, the rate of change in the gas-phase mass of $\alpha$ is a simple sum of sources and sinks,
\begin{equation} \label{eq:mdot_1}
    \dot{M}_\alpha =  y_\alpha\,\dot{M}_\star - \dot{M}_{\alpha,\,\text{remnants}} - \dot{M}_{\alpha,\,\text{outflows}}
\end{equation}
where $y_\alpha\,\dot{M}_{\star}$ describes \cc\ enrichment. 
In terms of the return mass fraction of stars $r$, the mass lost to remnants is $Z_\alpha\,\dot{M}_\star\,(1-r)$.  And, the outflows deplete mass at a rate $Z_\alpha \,\dot{M}_\star\,\eta$. (We assume the composition of outflows is the same as the interstellar medium.) With these definitions, Eq.~\ref{eq:mdot_1} is
\begin{equation}
    \dot{M}_\alpha= y_\alpha\,\dot{M}_\star - (1 + \eta - r)\,Z_\alpha\,\dot{M}_\star.
\end{equation}
Assuming an exponentially declining \sfh{} $\dot{M}_\star \propto e^{-t/\tau_{\rm sfh}}$, the equilibrium abundance is derived analytically by setting $\dot{Z}_\alpha=0$.
\begin{equation}\label{eq:z_eq}
    Z_\alpha^{\rm eq}(R) = \frac{y_\alpha}{1 + \eta(R) - r - \tau_\star / \tau_{\rm sfh}},
\end{equation}
where $\tau_{\star}$ is the star formation rate. ($\eta$ depends on $R$ to create a metallicity gradient. See also Section \ref{sec:vice} and Eq.~\ref{eq:eta_r}.)
In the special case of constant star formation, $\tau_{\rm sfh}\to\infty$, the denominator simplifies to $1+\eta-r$.

@@

@@

In Section \ref{sec:agb_results}, we will show that none of the four \agb\ yield sets (\cxi{}, \kx{}, \vxiii{}, \kxvi{}) produce enough C relative to this $\Ycc$ value. So, we introduce a normalization factor $\alpha_{\rm agb}$, which denotes a multiplicative scaling of $\Ycagb$ 
\begin{equation} \label{eq:alpha}
        \Ycagb \rightarrow \alpha_{\rm agb}\ \Ycagb.
\end{equation}
To parameterize the \agb\ contribution to C production, we define $f_{\rm agb}$ to be the fraction of C which comes from \agb\ stars. 
\begin{equation}\label{eq:f_agb}
    f_{\rm agb} \equiv \frac{\Ycagb(Z=\Zo)}{\Yct(Z=\Zo)},
\end{equation}
($f_{\rm agb}$ accounts for the multiplicative factor $\alpha_{\rm agb}$ as defined in Eq.~\ref{eq:alpha}).

@@

\subsection{CCSNe Carbon Yields}
With the total C yield of Eq.~\ref{eq:yc_inferred} and given an \agb\ C yield, we can derive an observationally-consistant \cc\ C yield.
\begin{subequations}
    \begin{gather}
        \Ycc(\Zo) = \Yct(\Zo) - \Ycagb(\Zo)\\
        \zeta^{\rm cc} \equiv \frac{d\Ycc(\Zo)}{dZ} = \frac{d\Yct(\Zo)}{dZ} - \frac{d\Ycagb(\Zo)}{dZ}
    \end{gather}
\end{subequations}

Furthermore, we can assert that the slope, $d \Yct / dZ$ at solar metallicity is consistant. We find from Fig.~\ref{fig:analytic} that

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{analytic.pdf}
    \caption[Reverse-Fit Yields]{Inferred high-mass star C yields as a function of metallicity. We assume equilibrium and $3\times {\rm C11}$ \agb\ yields (orange curve, see discussion in Section \ref{sec:equilibrium}). Blue points are the median value of $\Ycc$ for each bin in [Mg/H] with uncertainties based on the 16-84 percentile range.
    }
    \label{fig:analytic}
\end{figure}

\begin{equation}\label{eq:zeta}
    \Ycc = y_0 + \zeta\,(Z-\Zo) + \frac{2\,y_{\rm III}}{1 + Z/Z_{\rm III}},
\end{equation}
where $\zeta\approx0.001$ describes the metallicity dependence.

@@

Our model is as follows. The Galaxy is divided into 200 rings, each 100\,pc wide. Each ring has a separate stellar population and gas supply. We initially assume an inside-out \sfh{}, where the star formation surface density $\Sigma_\star$ is given by 
\begin{equation}\label{eq:inside_out}
    \dot{\Sigma}_\star \propto \left(1-e^{-t/\tau_{\rm rise}}\right) e^{-t/\tau_{\rm sfh}}.
\end{equation}
$\tau_\text{rise}=2$\,Gyr describes when the star formation rate reaches a maximum, and $\tau_{\rm sfh}$ describes the decay timescale of star formation as a function of radius $R$. \JJ~derives $\tau_{\rm sfh}(R)$ through analysis of four integral field spectroscopy surveys in \cite{sanches20}. At each $R$, the \sfh{} is normalized to match the stellar surface density gradient \citep{BHG16} and the total stellar mass reaches $5.17\times10^{10}\,\Mo$ \citep{LM15}. Star formation ends beyond a radius $R=15.5\,$kpc. 
The gas inflow is calculated to maintain our chosen \sfh{} for each radius and time, using an extension of a Kennicutt-Schmidt law \citep{kennicutt98},
\begin{equation}
\dot{\Sigma}_{\star} \propto 
\begin{cases}
    {\Sigma}_{\rm gas} & 2\times 10^7 \leq \Sigma_{\rm gas} \\ 
    {(\Sigma}_{\rm gas})^{3.6} & 5\times 10^6 \leq \Sigma_{\rm gas} < 2\times10^7 \\ 
    {(\Sigma}_{\rm gas})^{1.7} & \Sigma_{\rm gas} < 5\times10^6 \\ 
\end{cases}
\end{equation} 
where $\Sigma_{\rm gas}$ is measured in \Mo\,kpc$^{-2}$. 
The scaling of this relationship varies with time due to the redshift dependence of $\tau_\star$ in molecular gas observed by \citet{tacconi18}. We assume a \citet{kroupa01} \imf.

@@

Subgiants provide the ideal observational constraint to our model. Since first dredge up changes C surface abundances in evolved stars, these stars require model-dependent corrections to recover surface abundances \citep[e.g.][]{vincenzo+21}. On the other hand, gravitational settling can affect main sequence stellar abundances \citep[e.g.][]{souto19}. Subgiants have well-mixed envelopes, so gravitational settling is not as significant, and subgiants have not yet experienced first dredge up. We use a sample of \apogee\ DR17 stars \citep{apogee17} as selected in \citetjack. We only compare the models to the low-$\alpha$ sequence (representing the thin disk) unless otherwise specified.  The selection criteria and differences between the samples are described in more detail in Appendix~\ref{sec:jack}.

@@

C evolution proceeds as follows, (for a single zone)
\begin{enumerate}
    \item \cc\ initially dominate production. As $\Ycc$ has strong metallicity dependence, [C/Mg] increases with time. 
    \item \agb stars contribute delayed C, causing [C/Mg] to increase even faster with [Mg/H]. 
    \item[(iii)] [C/Mg] plateaus as C also approaches equilibrium. 
    \item[(iv)] [C/Mg] may decrease due to declining SFH or slightly negative yields from \about{1}\,\Mo stars.

@@

\begin{figure*}
\centering
\includegraphics{all_the_tracks.pdf}
\caption[Carbon Chemical Evolution Tracks]{
    Time evolution of gas-phase C abundances in our fiducial model.
    Each line represents a zone at a different galactic radii. The lines are coloured-coded by time. The top shows \caah\ and the bottom \caafe. We artificially restrict the axis range in each case to better see solar-annulus and present-day evolution.
}
\label{fig:c_evo}
\end{figure*}

@@

\subsection{High-mass stellar yields}\label{sec:results_highmass}
\cc, which likely make up the majority of C, set the overal C abundances. 
 and the \cc\ metallicity dependence $\zeta$ in Fig.~\ref{fig:first_models}. Fig.~\ref{fig:first_models} shows models with varying $\Ycc$ metallicity dependence. As the \caah~trend is approximated by the equilibrium C/Mg abundance with metallicity, the models with stronger metallicity dependence have a stronger slope in \caah. However, \caafe~is minimally affected by these changes since \cc\ occurs on much shorter timescales than \ia\ and \agb\ enrichment, and we only consider stars in a small metallicity range for \caafe.

\begin{figure*}
\includegraphics{zeta.pdf}

\caption[Median Stellar Abundance Trends]{
    Stellar abundance trends in our model, assuming metallicity independent $\Ycc$. Colored lines quantify the median [C/Mg] in bins of [Mg/H] for our four \agb\ yield models from the literature (see Section \ref{sec:agb}). Black points and grey dashes represent the median and standard deviations of [C/Mg] for each [Mg/H] bin in the \citetjack~sample. In the right panel, we show the trends only for stars where $-0.15\leq {\rm [Mg/H]}\leq -0.05$.
}
\label{fig:first_models}
\end{figure*}

\subsection{Low mass stellar yields}\label{sec:agb_results}

Fig.~\ref{fig:agb_sims} compares the \kx{}, \kxvi{}, \cxi{}, and \vxiii{} yield models. As the highest \agb\ yield at solar, \kx{} is only $\Ycagb=0.000585$, so $f_{\rm agb} \leq 0.12$ for all models. 
Massive stars dominate C production with unscaled yields.  
With the exception of \vxiii, the different \agb\ models only affect the overall metallicity dependence (in \caah). We leave a more detailed discussion of the effects of \agb\ models for Appendix~\ref{sec:oob_models} and use \cxi{} (OR ANALYTIC?) yields hereafter.

\begin{figure*}
\centering
\includegraphics{caah_agbs.pdf}
\includegraphics{caah_indifferent.pdf}

\caption[]{Similar to Fig.~\ref{fig:first_models}, except where each line corresponds to one of our four \agb\ yield sets (\cxi, \kx, \vxiii, \kxvi).}
\label{fig:agb_sims}
\end{figure*}

@@

FIGUREEE

\begin{figure}
\centering
\includegraphics{caafe_agbs.pdf}

\caption[]{Similar to Fig.~\ref{fig:first_models}, except where the \agb\ C contribution is varied from the fiducial ($f_{\rm agb}=0.2$) to 0.1 and 0.3 (see Eq.~\ref{eq:f_agb} and Section.~\ref{sec:agb_results}.}
\label{fig:caafe_agbs}
\end{figure}

\begin{figure}
\centering
\includegraphics{caafe_fagb.pdf}

\caption[]{Similar to Fig.~\ref{fig:first_models}, except where the \agb\ C contribution is varied from the fiducial ($f_{\rm agb}=0.2$) to 0.1 and 0.3 (see Eq.~\ref{eq:f_agb} and Section.~\ref{sec:agb_results}.}
\label{fig:f_agb}
\end{figure}

@@

In this section, we consider two modifications of our fiducial \sfh{}: \textit{lateburst} and \textit{earlyburst}.
Our lateburst model adds a Gaussian factor to the inside-out \sfh{},
\begin{equation}\label{eq:late-burst}
    \dot{\Sigma}_\text{lateburst} \propto \dot{\Sigma}_\text{inside-out} \left(1 + A\,e^{-(t-\tau_{\rm burst})^2/2\sigma^2_{\rm burst}} \right)
\end{equation}
where $A=1.5$ represents the amplitude of the birth, $\tau_\text{burst}=10.8$\,Gyr is the time where the burst is strongest, and $\sigma_\text{burst}=1$\,Gyr is the width of the burst.
Our earlyburst model is similar, but the burst is exponential and placed at $t_1=5$\,Gyr. 
\begin{equation}\label{eq:twoinfall}
    \dot{\Sigma}_{\rm earlyburst} \propto \dot{\Sigma}_{\rm inside-out} + 
\begin{cases}
    A\,e^{-(t-t_1)/\tau_{\rm burst}} & t_1 < t \\
      0 & t<t_1
\end{cases}
\end{equation}
where we take the burst duration, $\tau_{\rm burst}=1$\,Gyr.
This approximately corresponds to the Gaia-Encelidus merger, inducing higher star formation in the Milky Way \citep{spitoni21, bonaca20, helmi18}.

Fig.~\ref{fig:sfh_models} shows three models with these alternate \sfh{}. Changes to the \sfh{} leave \caah\ unchanged, but they do introduce slight variation in \caafe. Models with higher \agb{} fractions are more sensitive to variations in the \sfh{}. The late burst models result in [C/Mg] continuing to increase at low [Mg/Fe], but also introduce a dip not present in the data. Additionally, the early burst
reproduces the slight break between the low and high $\alpha$ sequences, but overshoots equilibrium more severely than the fiducial model. 
In general, any of these \sfh{}s are consistent with this model.

\begin{figure}
\centering
\includegraphics{caafe_sfh.pdf}

\caption{Similar to Fig.~\ref{fig:first_models} but comparing the fiducial model to alternate \sfh{}s (see Section  \ref{sec:sfh}).}

>>>>

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
\usepackage{anyfontsize}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

% Acronyms
\newcommand{\agb}{AGB}
\newcommand{\apogee}{APOGEE}
\newcommand{\cc}{CCSNe}
\newcommand{\gce}{GCE}
\newcommand{\ia}{SNe Ia}
\newcommand{\imf}{IMF}
\newcommand{\sfh}{SFH} % check where first defined
\newcommand{\ssp}{SSP}

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. A. Boyea et. al.]{
Daniel A. Boyea,$^{1}$\thanks{E-mail: boyea.2@osu.edu}
James W. Johnson,$^{1}$
Third Author$^{2,3}$
and Others$^{1,3}$
\\
% List of institutions
$^{1}$Department of Astronomy, the Ohio State University, 191 W. Woodruff, Columbus, OH 43210, USA
}

@@

% Abstract of the paper
\begin{abstract}
% context
C is an important element across astronomy. However, its origin remains poorly understood. 
% 
As stellar yields are highly uncertain, we aim to constrain the stellar yields of C through multi-zone Galactic chemical evolution models. 
% results
We use \apogee\ subgiants to constrain the model. 
% 
We find that \caafe\ is an emirical estimate of the delayed C sources, enabling us to estimate that \agb\ stars and \cc\ produce about 20\% and 80\% of C, respectively.  
The \caah\ trend instead represents the equilibrium abundances of C and Mg. 
We use the \caah\ trend to estimate the CCSNe C/Mg yield, determining that  $\Ycc/\Yoc = EQUATION$ when including \agb\ C. 
% misc
Our models are relatively independent of uniform scaling of yields and outflows, and alternate star formation histories. 
However, the stars which contribute to \agb\ C production and the \ia\ delay time distribution of Fe contribute uncertainties to our conclusions. 
% gas phase
While reliable gas-phase and low-metallicity measurments of C are challenging, we find that our model and a singlezone model with our recommended yields replicate the broad trends of \caah across different environments and metallicities.

@@

One of the primary uncertainties of GCE models are nucleosynthetic yields. Yield predictions -- the amounts of each chemical element stars produce --
are shaped by poorly understood processes, including mass loss, nuclear reaction rates, rotational mixing, convection, and explodability \citep{romano+10,KL14,ventura+13, LC18, emily+21}.
To better understand where C comes from and how it evolves, our aim is to use a combination of \apogee\ observations and multi-zone models to develop estimates of the required C yields to reproduce galactic trends.
\cite{james+23} examined similar \gce{} models of N (which is closely related to C), finding that trends in N and O\footnotemark{} are explained by the metallicity dependence of N/O yields. \citet{james+23} determine that \agb\ N abundances roughly dependence on metallicity (i.e. $y_{\rm N}/y_{\rm O} \propto Z$). 
Here, we extend their models to C, deriving similar constraints on C/Mg yields. We assess which yield prescriptions reproduce Galactic abundance trends while investigating the impact of \gce{} model assumptions, such as the star formation history (\sfh{}) and outflow mass loading.

@@

\begin{figure*}
    \centering
    \includegraphics{subgiants.pdf}
    \caption{The [C/Mg] ratio against [Mg/H] (top) and [Mg/Fe] (bottom) for the \citetjack~sample of \apogee{} subgiants. On the top, we plot high and low-$\alpha$ stars in blue and orange, using the separation defined in Equation \ref{eq:high_alpha} (the high and low-$\alpha$ stars are named for their high or low $\alpha$-element to Fe ratios, or in this case, Mg/Fe). On the bottom, we colour-code stars according to their [Mg/H] abundance.} \label{fig:subgiants}
\end{figure*}

@@

\begin{table}
	\centering
    \caption[]{Yields for the fiducial model (in units of \ssp~birth mass). See Section \ref{sec:agb} for the definition of \cxi.}
	\label{tab:fiducial_mod}

@@

Here, we assume a simple \textit{one-zone} chemical evolution model \cite[e.g.][]{tinsley80, pagel09, matteucci21}.  Newly produced metals are homogeneously and instantaneously mixed, so this model evolves as a single, spatially-uniform chemical zone.
We define $M_{X}$ to be the mass of element $X$ in the gas-phase, $\dot{M}_\star$ to be the star formation rate (in $\Mo\,{\rm yr}^{-1}$), and $\eta$ to be the mass loading factor $\eta\equiv\dot{M}_{\rm outflow}/\dot{M}_\star$ (representing the strength of outflows). A \ssp\ returns a fraction $r$ of their birth mass to the interstellar medium, due to ejected stellar envelopes.%
\footnote{$r\approx0.4$ for a \citealt{kroupa01} \imf.}
Here, we will use $\alpha$ as a representative $\alpha$-element, such as O or Mg. 
Given the \imf-averaged yield $y_\alpha$, the rate of change in the gas-phase mass of $\alpha$ is a simple sum of sources and sinks,
\begin{equation} \label{eq:mdot_1}
    \dot{M}_\alpha =  y_\alpha\,\dot{M}_\star - \dot{M}_{\alpha,\,\text{remnants}} - \dot{M}_{\alpha,\,\text{outflows}}
\end{equation}
where $y_\alpha\,\dot{M}_{\star}$ describes \cc\ enrichment. 
In terms of the return mass fraction of stars $r$, the mass lost to remnants is $Z_\alpha\,\dot{M}_\star\,(1-r)$.  And, the outflows deplete mass at a rate $Z_\alpha \,\dot{M}_\star\,\eta$. (We assume the composition of outflows is the same as the interstellar medium.) With these definitions, Eq.~\ref{eq:mdot_1} is
\begin{equation}
    \dot{M}_\alpha= y_\alpha\,\dot{M}_\star - (1 + \eta - r)\,Z_\alpha\,\dot{M}_\star.
\end{equation}
Assuming an exponentially declining star formation history (\sfh{}) $\dot{M}_\star \propto e^{-t/\tau_{\rm sfh}}$, the equilibrium abundance is derived analytically by setting $\dot{Z}_\alpha=0$.
\begin{equation}\label{eq:z_eq}
    Z_\alpha^{\rm eq}(R) = \frac{y_\alpha}{1 + \eta(R) - r - \tau_\star / \tau_{\rm sfh}},
\end{equation}
where $\tau_{\star}$ is the star formation rate. ($\eta$ depends on $R$ to create a metallicity gradient. See also Section \ref{sec:vice} and Eq.~\ref{eq:eta_r}.)
In the special case of constant star formation, $\tau_{\rm sfh}\to\infty$, the denominator simplifies to $1+\eta-r$.

@@

\begin{figure}
    \centering
    \includegraphics{analytic.pdf}
    \caption[]{Inferred high-mass star C yields as a function of metallicity. We assume equilibrium and $3\times {\rm C11}$ \agb\ yields (orange curve, see discussion in Section \ref{sec:equilibrium}). Blue points are the median value of $\Ycc$ for each bin in [Mg/H] with uncertainties based on the 16-84 percentile range.
    }
    \label{fig:analytic}
\end{figure}

@@

To parameterize the \agb\ contribution to C production, we define $f_{\rm agb}$ to be the fraction of C which comes from \agb\ stars. 
\begin{equation}\label{eq:f_agb}
    f_{\rm agb} \equiv \frac{\Ycagb(Z=\Zo)}{\Yct(Z=\Zo)},
\end{equation}
In Section \ref{sec:agb_results}, we will show that $f_{\rm agb} \approx 0.2$. 
None of the \agb\ yield sets (\cxi{}, \kx{}, \vxiii{}, \kxvi{}) produce enough C relative to our total $\Ycc$ values. So, we introduce a normalization factor $\alpha_{\rm agb}$, which denotes a multiplicative scaling of $\Ycagb$ 
\begin{equation} \label{eq:alpha}
        \Ycagb \rightarrow \alpha_{\rm agb}\ \Ycagb.
\end{equation}
Table~\ref{tab:alpha_agb} contains the required values of $\alpha_{\rm agb}$ to reach $f_{\rm agb}=0.2$, the yields at solar metallicity, and the metallicity for each \agb\ yield set.

@@

\subsection{CCSNe Carbon Yields}
With the total C yield of Eq.~\ref{eq:yc_inferred} and given an \agb\ C yield, we can derive an observationally-consistant \cc\ C yield.
\begin{subequations}
    \begin{gather}
        \Ycc(\Zo) = \Yct(\Zo) - \Ycagb(\Zo)\\
        \zeta^{\rm cc} \equiv \frac{d\Ycc(\Zo)}{dZ} = \frac{d\Yct(\Zo)}{dZ} - \frac{d\Ycagb(\Zo)}{dZ}
    \end{gather}
\end{subequations}
This method reduces the number of free parameters of the model and enables all models to match \caah\ trends in \apogee{} subgiants when considering \agb\ yields with different metallicity dependences and normalizations. 
Table~\ref{tab:alpha_agb} contains our adaptive (regression-derived) values of $\zeta^{\rm agb}$ and $\Ycagb(\Zo)$. Table~\ref{tab:alpha_agb} also shows $\alpha^{\rm agb}_{20}$, the value of $\alpha^{\rm agb}$ required with our adopted total yields (Eq.~\ref{eq:yc_inferred}) to reach 20\% \agb\ C.

Our full \cc\ C yield model adds a low-metallicity enhancement term, which is insignificant at solar metallicities (e.g. thin disk stars) but enables models in Section~\ref{sec:gas} to match low-metallicity environments.
All-together, 
\begin{equation}\label{eq:zeta}
    \Ycc = y_{\rm C, 0}^{\rm cc} + \zeta^{\rm cc}\,(Z-\Zo) + \frac{2\,y_{\rm III}}{1 + Z/Z_{\rm III}}.
\end{equation}

@@

Star formation is treaded as follows. The Galaxy is divided into 200 rings, each 100\,pc wide. Each ring has a separate stellar population and gas supply. We initially assume an inside-out \sfh{}, where the star formation surface density $\Sigma_\star$ is given by 
\begin{equation}\label{eq:inside_out}
    \dot{\Sigma}_\star \propto \left(1-e^{-t/\tau_{\rm rise}}\right) e^{-t/\tau_{\rm sfh}}.
\end{equation}
$\tau_\text{rise}=2$\,Gyr describes when the star formation rate reaches a maximum, and $\tau_{\rm sfh}$ describes the decay timescale of star formation as a function of radius $R$. \JJ~derives $\tau_{\rm sfh}(R)$ through analysis of four integral field spectroscopy surveys in \cite{sanches20}. At each $R$, the \sfh{} is normalized to match the stellar surface density gradient \citep{BHG16} and the total stellar mass reaches $5.17\times10^{10}\,\Mo$ \citep{LM15}. Star formation ends beyond a radius $R=15.5\,$kpc. 
The gas inflow is calculated to maintain the \sfh{} for each radius and time, using an extension of a Kennicutt-Schmidt law \citep{kennicutt98},
\begin{equation}
\dot{\Sigma}_{\star} \propto 
\begin{cases}
    {\Sigma}_{\rm gas} & 2\times 10^7 \leq \Sigma_{\rm gas} \\ 
    {(\Sigma}_{\rm gas})^{3.6} & 5\times 10^6 \leq \Sigma_{\rm gas} < 2\times10^7 \\ 
    {(\Sigma}_{\rm gas})^{1.7} & \Sigma_{\rm gas} < 5\times10^6 \\ 
\end{cases}
\end{equation} 
where $\Sigma_{\rm gas}$ is measured in \Mo\,kpc$^{-2}$. 
The scaling of this relationship varies with time due to the redshift dependence of $\tau_\star$ in molecular gas observed by \citet{tacconi18}. We assume a \citet{kroupa01} \imf.

@@

Subgiants provide the ideal observational constraint to our model. 
When a star enters the Red Giant Branch (RGB), material from the CNO-processed core is mixed with the envelope in first dredge up, enhancing N and depleting C \citep{iben67, vincenzo+21,KL14}. RGB stars thus require model-dependent corrections to recover surface abundances \citep[e.g.][]{vincenzo+21}. On the other hand, gravitational settling can affect main sequence stellar abundances \citep[e.g.][]{souto19}. Subgiants have well-mixed envelopes, so gravitational settling is not as significant, and subgiants have not yet experienced first dredge up. We use a sample of \apogee\ DR17 stars \citep{apogee17} as selected in \citetjack. We only compare the models to the low-$\alpha$ sequence (representing the thin disk) unless otherwise specified.  The selection criteria and differences between the samples are described in more detail in Appendix~\ref{sec:jack}.

@@

C evolution proceeds as follows, (for a single zone)
\begin{enumerate}
    \item \cc\ initially dominate production. As $\Ycc$ has strong metallicity dependence, [C/Mg] increases with time. 
    \item \agb\ stars contribute delayed C, causing [C/Mg] to increase even faster with [Mg/H]. 
    \item{} [C/Mg] plateaus as C also approaches equilibrium. 
    \item{} [C/Mg] may decrease due to declining SFH or slightly negative yields from \about{1}\,\Mo stars.

@@

\begin{figure*}
\centering
\includegraphics{all_the_tracks.pdf}
\caption[]{
    Time evolution of gas-phase C abundances in our fiducial model.
    Each line represents a zone at a different galactic radii. The lines are coloured-coded by time. The left shows \caah\ and the right \caafe. We artificially restrict the axis range in each case to better see solar-annulus and present-day evolution.
    <UPDATE, CHECK LEFT?RIGHT USAGE>
}
\label{fig:c_evo}
\end{figure*}

@@

\subsection{High-Mass Stellar Yields}\label{sec:results_highmass}

\cc, which likely make up the majority of C, set the overall C abundance trends.
 Fig.~\ref{fig:first_models} shows models with varying $\Ycc$ metallicity dependence. As the \caah~trend is approximated by equilibrium trends, the models with stronger metallicity dependence have a stronger slope in \caah. 
 However, \caafe~is minimally affected by these changes since \cc\ occurs on much shorter timescales than \ia\ and \agb\ enrichment. The only effects on \caafe, when considering the narrow metallicity slice, are because of either the slight change in equilibrium abundances, the imperfect evolution of the galaxy, or that the interstellar medium abundances are set by stars which were born at poorer metallicities. 
 Hence, \caah\ tells us about the total C yield with metallicity, which \caafe\ is independent of. If we know the \agb\ C yields, then with observed \caafe\ abundance trends, we can infer the \cc\ C yields with metallicity.

\begin{figure*}
\includegraphics{all_p1.pdf}

\caption[]{
    Stellar abundance trends in our model, assuming metallicity independent $\Ycc$. Colored lines quantify the median [C/Mg] in bins of [Mg/H] for our four \agb\ yield models from the literature (see Section \ref{sec:agb}). Black points and grey dashes represent the median and standard deviations of [C/Mg] for each [Mg/H] bin in the \citetjack~sample. In the right panel, we show the trends only for stars where $-0.15\leq {\rm [Mg/H]}\leq -0.05$.
}
\label{fig:first_models}
\end{figure*}

\subsection{Low-Mass Stellar Yields}\label{sec:agb_results}

\caafe\ is sensitive to the assumptions about delayed C from \agb\ stars. If no C comes from low-mass stars, then [C/Mg] would be independent of [Mg/Fe], only [Mg/H]. Instead, C shows strong trends  in \caafe\, independent of metallicity.

In Fig. <REF>, we first show the four C yield models (\kx, \kxvi, \cxi, \vxiii). For the most part, the \agb\ yield sets result in qualitatively similar predictions. \vxiii, however, does not reproduce solar trends as the model predicts strong C production at slightly above solar metallicities, resulting in a decreasing [C/Mg] with [Fe/Mg]. 
As all \agb\ models predict some low-mass C destruction, each model does predict a downturn in [C/Mg] as [Fe/Mg] increases. A recent burst in star formation may hide this downturn (see NEXT SECTION), but the C destruction is not supported directly by observations.

@@

\begin{figure*}
\centering
\includegraphics{all_p2.pdf}

\caption[]{Similar to Fig.~\ref{fig:first_models}, except where the \agb\ C contribution is varied from the fiducial ($f_{\rm agb}=0.2$) to 0.1 and 0.3 (see Eq.~\ref{eq:f_agb} and Section.~\ref{sec:agb_results}.}
\label{fig:caafe_agbs}
\end{figure*}

@@

In this section, we consider two modifications of our fiducial \sfh{}: \textit{lateburst} and \textit{earlyburst}.
Our lateburst model adds a Gaussian factor to the inside-out \sfh{},
\begin{equation}\label{eq:lateburst}
    \dot{\Sigma}_\text{lateburst} \propto \dot{\Sigma}_\text{inside-out} \left(1 + A\,e^{-(t-\tau_{\rm burst})^2/2\sigma^2_{\rm burst}} \right)
\end{equation}
where $A=1.5$ represents the amplitude of the birth, $\tau_\text{burst}=10.8$\,Gyr is the time where the burst is strongest, and $\sigma_\text{burst}=1$\,Gyr is the width of the burst.
Our earlyburst model instead >>>>>?????  at $t_1=5$\,Gyr. 
\begin{equation}\label{eq:twoinfall}
    \dot{\Sigma}_{\rm earlyburst} \propto A\,e^{-(t-t_1)/\tau_{\rm burst}} + 
\begin{cases}
    \left(1-e^{-t/\tau_\text{rise}}\right)e^{-t/\tau_\text{rise}} & t_1 < t \\
      0 & t<t_1
\end{cases}
\end{equation}
where we take the burst duration, $\tau_{\rm burst}=1$\,Gyr.
This approximately corresponds to the Gaia-Encelidus merger, inducing higher star formation in the Milky Way \citep{spitoni21, bonaca20, helmi18}.

Fig.~\ref{fig:sfh_models} shows three models with these alternate \sfh{}. Changes to the \sfh{} leave \caah\ unchanged, but they do introduce slight variation in \caafe. Models with higher \agb{} fractions are more sensitive to variations in the \sfh{}. The late burst models result in [C/Mg] continuing to increase at low [Mg/Fe], but also introduce a dip not present in the data. Additionally, the early-burst
reproduces the slight break between the low and high $\alpha$ sequences, but overshoots equilibrium more severely than the fiducial model. 
In general, any of these \sfh{}s are consistent with this model.
